---
title: "RIDB-Watchman-2019-data"
author: "Soyoung Park"
date: "6/19/2020"
---

#RIDB-Watchman project
##Watchman overview
*FacilityID: 232445 (unique ID)
*LegacyFacilityID: 70923	
*OrgFacilityID: 128
*ParentOrgID: AN370923
*ParentRecAreaID: 2994
*FacilityName: WATCHMAN CAMPGROUND
*FacilityUseFeeDescription: A \$10.00 service fee will apply if you change your reservation.  Cancellations more than 48 hours in advance of arrival date incur a \$10 cancellation fee. Within 48 hours of arrival, a \$10 cancellation fee and the first night's fee are charged.  "	Campground	Zion Group Site fees are \$50, \$90 or \$130 per night based on the campsite capacity. Check the minimum and maximum capacity for each group site. Each group sites minimum required occupants vary.
*FacilityLongitude: -112.986389	
*FacilityLatitude: 37.198611	

```{r load packages}
library(readr)
library(data.table)
require(psych)
require(dplyr)
require(tidyr)
require(reshape2)
require(lubridate)
```

```{r load facility and campsite info}
rm(list=ls())
#set working directory
setwd("~/Documents/GitHub/RIDB-Watchman")

#load data
#facilities <- fread("~/Downloads/reservations2019.csv") #campsite data from Will. Size 1.2GB
#head(facilities)
#glimpse(facilities)
#Watchman <- facilities[facilities$facilityid==232445,] #watchman campground info
#Watchman <- facilities[facilities$park=="WATCHMAN CAMPGROUND",] #watchman campground info
#fwrite(Watchman, file = "~/Documents/GitHub/RIDB-Watchman/Watchman2019.csv")

watchman <- fread("~/Documents/GitHub/RIDB-Watchman/Watchman2019.csv")

campsites <- fread("~/Documents/GitHub/RIDB-Watchman/data/Campsites_API_v1.csv") #campsite characteristics
campsites <- campsites[campsites$FacilityID==232445,]
glimpse(campsites)
unique(campsites$CampsiteType)
```

```{r load reservation data}
# reservation <- fread("~/Documents/GitHub/RIDB-Watchman/reservations2018.csv") #2018 RIDB dataset
# colnames(reservation) #column names - FacilityID exists
# 
# #subset data just for Watchman
# reservation <- reservation[reservation$FacilityID==232445,]
# glimpse(reservation)
# unique(reservation$Park) #confirming that the data only has Watchman as the campground
# 
# fwrite(reservation, file = "~/Documents/GitHub/RIDB-Watchman/Watchman.csv")
```

```{r load Watchman reservation data}
rm(list=ls())

reservation <- watchman
unique(reservation$parentlocation) # Zion National Park
unique(reservation$park) # Watchman
count(reservation, 'hisoricalreservationid') # 24683 obersvations
count(distinct(reservation)) # 24683 obersvations

camp <- reservation[grepl("Overnight", reservation[["usetype"]]), ] #24676 observations. Removed 7 data points

unique(camp$sitetype) #"TENT ONLY NONELECTRIC"/"STANDARD ELECTRIC"/"WALK TO"/"GROUP TENT ONLY AREA NONELECTRIC"/"MANAGEMENT"                      
length(unique(camp$productid)) #179
```

```{r How many dates do people book in advance}
head(camp)

#calculate difference between the reservation date and the start date
camp$bookingWindow <- as.Date(as.character(camp$startdate), format="%Y-%m-%d") - as.Date(as.character(camp$orderdate), format="%Y-%m-%d")
camp$bookingWindow <- as.numeric(camp$bookingWindow)

summary(camp$bookingWindow) #there are negative booking windows

camp[camp$bookingWindow<0 ,] #there are 737 negative date differences between the reservation and the start date. This could be system error as most of them have order date of 2019-10-01 00:00:00 and start date in 2018. Only a few had the order date same as the end date.

camp <- camp[camp$bookingWindow>-1,] #remove 737 negative booking window (data reduced from 24676 to 23939)

summary(camp$bookingWindow)
#mean 102 days, median 121 days, min 0, max 331
```

```{r length of stay}
#calculate difference between the end date and the start date
camp$lengthStay <- as.Date(as.character(camp$enddate), format="%Y-%m-%d") - as.Date(as.character(camp$startdate), format="%Y-%m-%d")

camp$lengthStay <- as.numeric(camp$lengthStay) #number of nights stayed
camp[camp$lengthStay<0 ,] # 14 observations with negative length of stay.
camp <- camp[camp$lengthStay>-1,] #remove 14 negative booking window (data reduced from 23939 to 23925)
summary(camp$lengthStay) 
#mean 0 days, max 360 days <- the dataset is a little wierd!
glimpse(camp)
```

```{r merge with campsite info}
campsites$EntityID <- as.integer(campsites$CampsiteID)
camp <- full_join(camp, campsites)

unique(camp$CampsiteName) #179 campistes.
count(camp[camp$CampsiteName=="",]) #1265 reservations do not have a Campsite Name associated with it. We delete these.

camp <- camp[!is.na(camp$CampsiteName),] #reduced to 27248 observations
```

```{r average booking window per campsite}
camp <- camp %>%
  group_by(CampsiteName) %>%
  mutate(avgBookingWindow = mean(bookingWindow)) 

summary(camp$avgBookingWindow)

fwrite(camp, file = "~/Documents/GitHub/RIDB-Watchman/WatchmanDataset.csv")

colnames(camp)

bookingwindow <- select(camp, OrgID, RegionCode, RegionDescription, ParentLocationID, ParentLocation, FacilityID, FacilityZIP, LegacyFacilityID, Park, SiteType, UseType, CampsiteID, CampsiteName, CampsiteType, TypeOfUse, Loop, CampsiteAccessible, CampsiteLongitude, CampsiteLatitude, avgBookingWindow)

fwrite(camp, file = "~/Documents/GitHub/RIDB-Watchman/WatchmanDataset_Campsite_level.csv")
```


