---
title: "RIDB-Watchman"
author: "Soyoung Park"
date: "3/13/2020"
---

#RIDB-Watchman project
##Watchman overview
*FacilityID: 232445 (unique ID)
*LegacyFacilityID: 70923	
*OrgFacilityID: 128
*ParentOrgID: AN370923
*ParentRecAreaID: 2994
*FacilityName: WATCHMAN CAMPGROUND
*FacilityUseFeeDescription: A \$10.00 service fee will apply if you change your reservation.  Cancellations more than 48 hours in advance of arrival date incur a \$10 cancellation fee. Within 48 hours of arrival, a \$10 cancellation fee and the first night's fee are charged.  "	Campground	Zion Group Site fees are \$50, \$90 or \$130 per night based on the campsite capacity. Check the minimum and maximum capacity for each group site. Each group sites minimum required occupants vary.
*FacilityLongitude: -112.986389	
*FacilityLatitude: 37.198611	
*LastUpdatedDate: 3/13/20

```{r load packages}
library(readr)
library(data.table)
require(psych)
require(dplyr)
require(tidyr)
require(reshape2)
require(lubridate)
```

```{r load facility and campsite info}
#set working directory
setwd("~/Documents/GitHub/RIDB-Watchman")

#load data
facilities <- fread("~/Documents/GitHub/RIDB-Watchman/RIDBFullExport_V1_CSV/Facilities_API_v1.csv") #campsite characteristics
glimpse(facilities)
Watchman <- facilities[facilities$FacilityName=="WATCHMAN CAMPGROUND",] #watchman campground info
rm(facilities)

campsites <- fread("~/Documents/GitHub/RIDB-Watchman/RIDBFullExport_V1_CSV/Campsites_API_v1.csv") #campsite characteristics
campsites <- campsites[campsites$FacilityID==232445,]
glimpse(campsites)
unique(campsites$CampsiteType)
```

```{r load reservation data}
# reservation <- fread("~/Documents/GitHub/RIDB-Watchman/reservations2018.csv") #2018 RIDB dataset
# colnames(reservation) #column names - FacilityID exists
# 
# #subset data just for Watchman
# reservation <- reservation[reservation$FacilityID==232445,]
# glimpse(reservation)
# unique(reservation$Park) #confirming that the data only has Watchman as the campground
# 
# fwrite(reservation, file = "~/Documents/GitHub/RIDB-Watchman/Watchman.csv")
```

```{r load Watchman reservation data}
rm(list=ls())

reservation <- fread("~/Documents/GitHub/RIDB-Watchman/Watchman.csv")
unique(reservation$RegionDescription) # Zion National Park
unique(reservation$Park) # Watchman
count(reservation, 'HisoricalReservationID') # 28522 obersvations
count(distinct(reservation)) # 28522 obersvations

camp <- reservation[grepl("Overnight", reservation[["UseType"]]), ] # all are campsites

unique(camp$SiteType)
unique(camp$EntityID) #189 <- this is the CampsiteID
unique(camp$ProductID) #189
```

```{r How many dates do people book in advance}
head(camp)

#calculate difference between the reservation date and the start date
camp$bookingWindow <- as.Date(as.character(camp$StartDate), format="%Y-%m-%d") - as.Date(as.character(camp$OrderDate), format="%Y-%m-%d")
camp$bookingWindow <- as.numeric(camp$bookingWindow)

summary(camp$bookingWindow) #there are negative booking windows

camp[camp$bookingWindow<0 ,] #there are 9 negative date differences between the reservation and the start date. This could be system error.

camp <- camp[camp$bookingWindow>-1,] #remove 9 negative booking window (data reduced from 28522 to 28513)
summary(camp$bookingWindow)
#mean 90 days, median 85 days, min 0, max 357
```

```{r length of stay}
#calculate difference between the end date and the start date
camp$lengthStay <- as.Date(as.character(camp$EndDate), format="%Y-%m-%d") - as.Date(as.character(camp$StartDate), format="%Y-%m-%d")

camp$lengthStay <- as.numeric(camp$lengthStay) #number of nights stayed
summary(camp$lengthStay) # no negative length of stay
#mean 2 days, max 16 days

glimpse(camp)
```

```{r merge with campsite info}
campsites$EntityID <- as.integer(campsites$CampsiteID)
camp <- full_join(camp, campsites)

unique(camp$CampsiteName) #179 campistes.
count(camp[camp$CampsiteName=="",]) #1265 reservations do not have a Campsite Name associated with it. We delete these.

camp <- camp[!is.na(camp$CampsiteName),] #reduced to 27248 observations
```

```{r average booking window per campsite}
camp <- camp %>%
  group_by(CampsiteName) %>%
  mutate(avgBookingWindow = mean(bookingWindow)) 

summary(camp$avgBookingWindow)

fwrite(camp, file = "~/Documents/GitHub/RIDB-Watchman/WatchmanDataset.csv")

colnames(camp)

bookingwindow <- select(camp, OrgID, RegionCode, RegionDescription, ParentLocationID, ParentLocation, FacilityID, FacilityZIP, LegacyFacilityID, Park, SiteType, UseType, CampsiteID, CampsiteName, CampsiteType, TypeOfUse, Loop, CampsiteAccessible, CampsiteLongitude, CampsiteLatitude, avgBookingWindow)

bookingwindow <- distinct(bookingwindow)

fwrite(bookingwindow, file = "~/Documents/GitHub/RIDB-Watchman/WatchmanDataset_Campsite_level.csv")

bookingwindow_campsite <- select(bookingwindow, -SiteType)
bookingwindow_campsite <- distinct(bookingwindow_campsite)

fwrite(bookingwindow_campsite, file = "~/Documents/GitHub/RIDB-Watchman/WatchmanDataset_Campsite_level_no_siteType.csv")
```


